---
title: "Assignment for AI4PH - Developing and Deploying Transparent and Reproducible Algorithms for Public Health"
output:  
  bookdown::html_document2:
    toc: true
    number_sections: true
    toc_float: true
    fig_caption: yes
    global_numbering: true 
---

# Introduction

In this assignment, you will validate the `stroke` model we developed in class using a different dataset: `valid_data.rds`. You will run into issues using this dataset as it is because this is a raw dataset without data harmonization, which means that some variables in this dataset are different from the harmonized dataset we used to train and evaluate the model. Your job here is to harmonize the validation data so that it's in the same format as the example data we used in class. You can refer to `train_data_variables.csv` to see the format in the harmonized train data.

# Load useful packages

```{r, message=FALSE, results='hide'}
# data manipulation and building and validating models
library(tidyverse)

# models building and evaluation
library(tidymodels)

# model implementation
library(plumber)
```

# Load data

For this validation, we'll use a different and uncurated subset of the `stroke` dataset. We first take a look at the data and you'll see that it isn't in the same format as shown in `train_data_variables.csv`.

```{r}
# load the data
valid_data <- readRDS("valid_data.rds")
```

```{r}
# show sample size and number of predictors
dim(valid_data)
```

```{r}
# show the first 6 rows
head(valid_data) 
```

```{r}
# show some information about the dataset
str(valid_data)
```

```{r}
# show the summary information about the dataset
summary(valid_data)
```

# Try the saved model/workflow on the validation data as it is

It is expected to have some errors.

```{r}
# read in the saved workflow object
workflow <- readRDS("stroke_lr_workflow.rds")
```

```{r}
# predict on the validation data
prediction_class <- workflow %>% 
  predict(new_data = valid_data,
          type = 'class')
```

# Harmonize the validation data - add your harmonization steps in the below cell

Compare the data dictionary of the train data (`train_data_variables.csv`) with the summary info of the validation data, complete the data harmonization below.

```{r}
# make a copy of the valid data
harmonized_valid_data <- valid_data

# Add your harmonization steps here

```

```{r}
# check and make sure the variables are now in the correct format.
summary(harmonized_valid_data)
```

## verify that now the workflow works on the harmonized validation data

```{r}
prediction_class <- workflow %>% 
  predict(new_data = harmonized_valid_data,
          type = 'class')

head(prediction_class)
```

```{r}
prediction_prob <- workflow %>% 
  predict(new_data = harmonized_valid_data,
          type = 'prob')

head(prediction_prob)
```

```{r}
# Combine test data with predictions
pred_results <- harmonized_valid_data %>% 
  bind_cols(prediction_class, prediction_prob)

head(pred_results %>% select(stroke, `.pred_class`, `.pred_0`, `.pred_1`))
```

```{r}
# model performance on the harmonized validation data set
custom_metrics <- metric_set(sens, spec, roc_auc)

custom_metrics(pred_results, truth = stroke, estimate = .pred_class, .pred_0)

pred_results %>% roc_curve(stroke, .pred_0) %>% autoplot()
```


## Save the harmonized data set

```{r}
saveRDS(harmonized_valid_data, "harmonized_valid_data.rds")
```

# Implement the model using `plumber`

[`Plumber`](https://www.rplumber.io/) allows you to create a web API by merely decorating your existing R source code with roxygen2-like comments.

To run `plumber`, you will need a separate `.R` file.

```{r}
# 'stroke_lr_plumber.R' is the location of the file shown above
pr("stroke_lr_plumber.R") %>%
  pr_run(port=8000)
```

To test the website you generated, input the below examples, and check the output:

```{r}
# Example 1
ind <- 1

t(pred_results[ind, 1:10])

cat(paste("===============", "\n"))
cat(paste("The observed class:", pred_results$.pred_class[ind], "\n"))
cat(paste("The expected predicted class:", pred_results$.pred_class[ind], "\n"))
cat(paste("The expected probability for class 0:", round(pred_results$.pred_0[ind], 3), "\n"))
cat(paste("The expected probability for class 1:", round(pred_results$.pred_1[ind], 3)))
```

```{r}
# Example 2
ind <- 100

t(pred_results[ind, 1:10])

cat(paste("===============", "\n"))
cat(paste("The observed class:", pred_results$.pred_class[ind], "\n"))
cat(paste("The expected predicted class:", pred_results$.pred_class[ind], "\n"))
cat(paste("The expected probability for class 0:", round(pred_results$.pred_0[ind], 3), "\n"))
cat(paste("The expected probability for class 1:", round(pred_results$.pred_1[ind], 3)))
```

```{r}
# Example 3
ind <- 465

t(pred_results[ind, 1:10])

cat(paste("===============", "\n"))
cat(paste("The observed class:", pred_results$.pred_class[ind], "\n"))
cat(paste("The expected predicted class:", pred_results$.pred_class[ind], "\n"))
cat(paste("The expected probability for class 0:", round(pred_results$.pred_0[ind], 3), "\n"))
cat(paste("The expected probability for class 1:", round(pred_results$.pred_1[ind], 3)))
```